### 2018-04-04 paper.Rmd
SM comments on viz section

Yet there are some challenges working with CHILDES, both for students and for advanced users. For beginners with a lower level of technical experience, mastering the system can be tricky. The CHILDES ecosystem uses a special file format (CHAT), which is stored in a plain text format but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed via by a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting to navigate for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

@@

childes-db offers a collection of browser-based interactive visualizations that enable data exploration through easy-to-use interfaces that don’t require any programming knowledge. These tools are intended for casual users who seek to extract useful developmental indices from CHILDES without any technical overhead. The visualizations available through childes-db are the following: Frequency Counts (lexical statistics), Derived Measures (lexical diversity), and Population Viewer (corpora statistics). All three of these visualizations were built using Shiny, an R package that enables easy building of web applications in R. For each visualization, a user specifies a subset of the data (in particular the collection, corpus, child, and speaker) and is also able to restrict the age range of the child with slider bars. Any number of corpora, children, or speakers may be selected for analysis, allowing for aggregate counts across a whole range of data in the CHILDES system, a task that could not have been easily done using previous methods. Along with the visualizations, childes-db makes the raw data itself available for download under a separate “Table” tab; this allows researchers to further explore the underlying data from the plots.

@@

The Frequency Counts app allows users to enter a word (or words) in a text box and see the frequency (in parts per million) in which this word was uttered across a child's age for a particular subset of the data, whether it is an individual child in the Brown corpus or all Spanish speaking children and their mothers. The ability to view the caregiver’s usage of a term alongside a child’s assists researchers in better understanding the relationship between input and uptake. Users can also plot word trajectories alongside each other (such as “mom” and “dad” or “front” and “back”) for convenient cross-word comparisons.

@@

There are a number of ways to measure the lexical diversity of an individual's productive speech. Some of these measures, including MLU-w (mean length of utterance in words), TTR (type-token ratio), MTLD (measure of textual lexical diversity), and voc-d are available in the Derived Measures app, which plots these measures across age for a given subset of data, again specified by collection, corpora, children, and speakers. Because of the underlying gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). In the English language, these two measures have been shown to be highly correlated. As with the Frequency Counts app, caregiver’s lexical diversity measures can be plotted alongside children's.

### Population Viewer

The population viewer is a tool that helps users understand the relative sizes of various corpora by displaying the total number of utterances or tokens in the available corpora at particular ages. In many instances the size of a given corpus may inform researchers’ decisions as to what is the best target dataset for analysis.

@@

Although the interactive analysis tools described above cover some of the most common use cases of CHILDES data, researchers interested in more detailed and flexible analyses will want to interface directly with the data in childes-db. Making use of the `R` programming language [@R-base], we provide the `childesr` package to help researchers accomplish this task. `R` is an open-source, extensible statistical computing environment that is rapidly growing in popularity across fields and is increasing in use in child language research [e.g. @song2015; @norrman2015]. The `childesr` package abstracts away the details of connecting to the database. Users can take advantage of the tools developed in the popular `dplyr` package [@R-dplyr], which make manipulating large datasets quick and easy. We describe the commands that the package provides and then give several worked examples of using the package for analyses.

@@

The `childesr` package is easily installed via CRAN, the comprehensive R archive network. To install, simply type: `install.packages("childesr")`. After installation, users have access to data loading functions that retrieve a variety of subsets of the data:

@@

One common use of the CHILDES corpora is estimating the frequency with which children hear different words. These frequency estimates are used both in the development of theory [e.g. frequent words are learned earlier; @goodman2008], and in the construction of age-appropriate experimental stimuli. One benefit of the `childesdb` interface is that it allows for easy analysis of how the frequencies of words change over development. Many of our theories in which children learn the structure of language from its statistical properties implicitly assume that these statistics are *stationary*, i.e. unchanging over development [e.g., @saffran1996]. However a number of recent analysis show that the frequencies with which infants encounter both linguistic and visual properties of their environment may change dramatically over development [@fausey2016], and these changing distributions may produce similarly dramatic changes in the ease or difficulty with which these regularities can be learned [@elman1993].

To demonstrate how one might discover such a non-stationary, we take as a case study the frequency with which children hear the color words of English (e.g. "blue", "green"). Color words tend to be learned relatively late by children, potentially in part due to the abstractness of the meanings to which they refer [see @wagner2013]. However, within the set of color words, the frequency with which these words are heard predicts a significant fraction of the variance in their order of acquisition [@yurovsky2015]. But are these frequencies stationary--e.g. do children hear "blue" as often at 12 months as they do at 24 months? We answer this question in two ways--first using the childesb shiny apps, and then using the `childesr` package.

#### Using shiny

To investigate whether the frequency of color words is stationary over development, a user can navigate to the Frequency app, and enter a set of color words into the `Word` selector separated by a comma: here "blue, red, green." Because the question of interest is about the frequency of words in the input (rather than produced by children), the `Speaker` field can be set to reflect this. In this example we select "Mother" because, across corpora in CHIDLES, the speaker tagged with the Mother role tends to produce most of the words that children hear. Because children learn most of their basic color words by the age of 5, the age range 1-5 years is a reaonable choice for `Ages to include`. The results of these selections are shown in Figure \ref{fig:color-shiny}. We can also create a hyperlink to store these set of choices so that we can share these results with others, or with ourselves in the future, byby clicking on the `Bookmark` button in the bottom left corner.

From this figure, it seems likely that children hear "blue" more frequently early in development, but the trajectories of the "red" and "green" are less clear. We also do not have a good sense of the errors of these measurements, are limited to just a few colors at a time before the plot becomes too busy, and cannot combine frequencies across speakers. To perform this analysis in a more compelling and complete way, a user can use the `childesr` interface to get more direct access to the CHILDES corpora.

```{r color-shiny, fig.width = 10, fig.height = 6, fig.cap = "An example of using the Frequency shiny app to explore how children's color input changes over development"}
knitr::include_graphics("images/color-shiny.png")
```

#### Using childesr

Figure \ref{fig:color-plot}.

```{r color-plot, fig.height = 5, fig.width = 10, fig.cap ="Color frequency as a function of age. Points represent means across transcripts, error bars represent 95% confidence intervals computed by nonparametric bootstrap", cache = TRUE, message = FALSE, warning = FALSE}
colors <- c("black", "white", "red", "green", "yellow", "blue", "brown", 
             "orange", "pink", "purple")

>>>>

Yet there are some challenges working with CHILDES, both for students and for advanced users. For beginners with a lower level of technical experience, mastering the system can be tricky. The CHILDES ecosystem uses a special file format (CHAT), which is stored in a plain text format but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed using a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting to navigate for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

@@

childes-db offers a collection of browser-based interactive visualizations that enable data exploration through easy-to-use interfaces that do not require any programming knowledge. These tools are intended for casual users who seek to extract useful developmental indices from CHILDES without any technical overhead. The visualizations available through childes-db are the following: Frequency Counts (lexical statistics), Derived Measures (lexical diversity), and Population Viewer (corpora statistics). All three of these visualizations were built using Shiny, an R package that enables easy building of web applications in R. For each visualization, a user specifies a subset of the data (in particular the collection, corpus, child, and speaker) and is also able to restrict the age range of the child with slider bars. Any number of corpora, children, or speakers may be selected for analysis, allowing for aggregate counts across a whole range of data in the CHILDES system, a task that could not have been easily done using previous methods. Along with the visualizations, childes-db makes the raw data itself available for download under a separate “Table” tab; this allows researchers to further explore the underlying data from the plots.

@@

<!-- sm: intro sentence here for the role of frequency in CLA research  --> 
The Frequency Counts app allows users to enter a word (or words) in a text box and see the frequency (in parts per million) in which this word was uttered across a child's age for a particular subset of the data, whether it is an individual child in the Brown corpus or all Spanish speaking children and their mothers. <!-- sm: "whether..." is awk.  -->
The ability to view the caregiver’s usage of a term alongside a child’s assists researchers in better understanding the relationship between input and uptake. Users can also plot word trajectories alongside each other (such as “mom” and “dad” or “front” and “back”) for convenient cross-word comparisons.

@@

<!-- sm: intro sentence here for the role of developmental metrics in CLA research and in downstream applications like SLP--> 
There are a number of ways to measure the lexical diversity of an individual's productive speech. <!-- sm: different intro; MLU isn't a lexical diversity measure -->
Some of these measures, including MLU-w (mean length of utterance in words), TTR (type-token ratio), MTLD (measure of textual lexical diversity), and voc-d are available in the Derived Measures app, which plots these measures across age for a given subset of data, again specified by collection, corpora, children, and speakers. <!-- sm: MLU is not a measure of lexical diversity-->
Because of the underlying gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). <!-- sm: I think we should go ahead and compute this for transcript with MOR tiers? Given that this was one of the first uses, I don't think Brian will like the 'underlying gaps' explanation-->
In the English language, these two measures have been shown to be highly correlated. <!-- sm: citation needed -->
As with the Frequency Counts app, caregiver’s lexical diversity measures can be plotted alongside those of children.

### Population Viewer
The population viewer is a tool that helps users understand the relative sizes and temporal extents of various corpora by displaying the total number of utterances or tokens in the available corpora at particular ages. In many instances the size of a given corpus may inform researchers’ decisions as to what is the best target dataset for analysis. <!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

@@

Although the interactive analysis tools described above cover some of the most common use cases of CHILDES data, researchers interested in more detailed and flexible analyses will want to interface directly with the data in childes-db. Making use of the `R` programming language [@R-base], we provide the `childesr` package to help researchers accomplish this task. `R` is an open-source, extensible statistical computing environment that is rapidly growing in popularity across fields and is increasing in use in child language research [e.g. @song2015; @norrman2015]. The `childesr` package abstracts away the details of connecting to <!-- sm: and querying-->  the database. Users can take advantage of the tools developed in the popular `dplyr` package [@R-dplyr], which make manipulating large datasets quick and easy. We describe the commands that the package provides and then give several worked examples of using the package for analyses.

@@

The `childesr` package is easily installed via CRAN, the comprehensive R archive network. To install, simply type: `install.packages("childesr")`. After installation, users have access to data loading functions that retrieve a variety of subsets <!-- sm: entities? variety of subsets sounds non-hierarchical-->  of the data:

@@

One common use of the CHILDES corpora is estimating the frequency with which children hear different words. These frequency estimates are used both in the development of theory (e.g. frequent words are learned earlier; @goodman2008), and in the construction of age-appropriate experimental stimuli. One benefit of the childes-db interface is that it allows for easy analysis of how the frequency of words changes over development. Many of our theories in which children learn the structure of language from its statistical properties implicitly assume that these statistics are *stationary*, i.e. unchanging over development (e.g., @saffran1996). However a number of recent analysis show that the frequencies with which infants encounter both linguistic and visual properties of their environment may change dramattically over development [@fausey2016], and these changing distributions may produce similarly dramatic changes in the ease or difficulty with which these regularities can be learned [@elman1993].

To demonstrate how one might discover such a non-staionarity, we take as a case study the frequency with which children hear the color words of English (e.g. "blue", "green"). Color words tend to be learned relatively late by children, potentially in part due to the abstractness of the meanings to which they refer (see @wagner2013). However, within the set of color words, the frequency with which these words are heard predicts a significant fraction of the variance in their order of acquisition [@yurovsky2015]. But are these frequencies stationary--e.g. do children hear "blue" as often at 12 months as they do at 24 months? We answer this question in two ways--first using the childesb shiny apps, and then using the `childesr` package.

```{r}
colors <- c("black", "white", "red", "green", "yellow", "blue", "brown", 
             "orange", "pink", "purple")
